### 2024-05-02 index.qmd
Init asymmetric homophily compartment paper

---
title: Epidemiological dynamics in idealized metapopulations defined by asymmetric homophily
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    cite-method: biblatex
    biblatexoptions: sorting=none, style=numeric-comp
    pdf-engine: latexmk
    pdf-engine-opt: -xelatex
authors:
  - name: Matthew Adam Turner
    affiliation: Stanford University Doerr School of Sustainability
    roles: writing, model development
    corresponding: true
    email: maturner@stanford.edu
  - name: Izabela Rezende
  - name: Benjamin Singer
  - name: Hannah S. Derwanz
  - name: James Holland Jones
    affiliation: Stanford University Doerr School of Sustainability
    roles: writing
    corresponding: true
bibliography: ../library.bib
classoption: abstract
keep-tex: true
abstract: |
  This is the abstract
geometry:
  - margin=1.1in
  - letterpaper
biblio-title: References
---

## Introduction

Due to rapidly increasing population size and density in a changing climate, the interdisciplinary science of pandemic preparedness is an increasing priority [@Rodo2021; @Adashi2022]. To complement our biological understanding and preventative tools for mitigating the risk of pandemics, social science is essential to understand how social structure and behavior determines how disease and disease prevention behaviors (such as vaccination[@Amin2017; @Omer2021]) are adopted by individuals in different group-structured populations [@Buckee2021; @Epstein2008; @SmaldinoJones2021]. Oftentimes in epidemiology, groups are defined by any different group partitions can be relevant to epidemiology, including age (RSV and Polio refs), occupation (frontline/remote refs), and, of course, political party [@VanBavel2020]. To understand how social structure affects . Using this we can infer the effect of different vaccine distribution allotments between the two groups that could possibly better protect one group or the other, or meet other optimal criteria. Computational analyses of metapopulation models can be used to create *in silico* experiments that would be impossible to implement in the real world, for instance to predict how different vaccination distribution strategies would best prevent transmission dynamics [@Singer2022].

Whether a new viral mutation or a cultural adaptation, beneficial *innovations* diffuse differently through different metapopulation structures. Turner and colleagues (2023) showed that novel innovations with higher fitness than an alternative best diffused when first introduced to a model minority group compared to a majority group (in terms of population size), in a metapopulation with a single other majority group [@Turner2023], and diffusion was further optimized when majority groups moderately preferred in-group interactions. Preference for in- versus out-group interactions is known as *homophily* [@McPherson2001], and introduced to a simple model metapopulation composed of a minority and majority group (by population size), where there exists a lower-fitness alternative, fitness-biased social learning.

Simple idealized models like we develop here serve as a theoretical ground truth against which we can better understand real-world complexities. For instance, schistosomaiasis hotspots can be predicted with some accuracy based on data from initial outbreak [@Singer2024], but metapopulation structure may have an underlying effect on the emergence of hotspots that, when accounted for, could increase predictive power of data-driven statistical models. Metapopulation structure with asymmetric homophily is not yet rigorously understood in an idealized theoretical setting, despite the fact that contact structure can be observed to affect transmission dynamics of sexually transmitted disease [@Morris1991], the spread of antibiotic-resistant bacteria across hospital wards [@Shapiro2020], the emergence and diffusion of drug-resistant malaria [@Ariey2003], the spread of schistosomiasis [@Ciddio2017], and provide disease reservoirs that drove the spread of tuberculosis and COVID-19 between some prison and community populations [@Sacchi2015; @Ryckman2021]. Contact structure has been rigorously shown to shape epidemiological dynamics [@Arthur2017; @Großmann2021], but not in settings where social networks are defined by asymmetric homophily, specifically. For data-driven applications, asymmetric homophily measurements may be easier or more reliable compared to the network structure itself.

To understand the effect of asymmetric homophily levels on epidemiological dynamics, we developed a computational model where metapopulations are composed of two groups, a minority and majority group defined by population size, where group members interact more or less frequently depending on group membership and group-level homophily levels, and whether interactions are structured by a social network (more like in a sexually-transmitted disease) or by random asymmetric homophilous mixing within and between groups (like in an airborne disease), adapting the models of Turner et al (2023) and Turner, Reynolds, and Jones (2024) for epidemiology by augmenting the social learning dynamic with SIR epidemiological dynamics.

Models like this enable us to rigorously understand how one important mechanism affects the course of disease and optimal intervention strategy. The cost of rigor, at this stage, is the exclusion of other important details that can affect disease or cultural transmission, including individual-level general health, access to clean running water, working conditions, and social factors such as asymmetric outgroup aversion levels between groups, which can undermine the adoption of adaptive behaviors [@EhretConstantino2022]. However, this initial mechanistic model ideally sets in motion a series of follow-on studies to better understand which of these additional factors alter model predictions to what extent. Model refinements and specializations can be added for specific epidemiological contexts, different scenarios and intervention strategies can be tested computationally for optimal policies given available information. More shortcomings, refinements, and specializations can follow that, with the process repeating as necessary. This work complements a broader effort across evolutionary sciences to understand the interplay between social structure and collective co-adaptation among humans, non-humans, and pathogens [@Bak-Coleman2021; @GalesicEtAl2023].

## Model

## Analysis

## Discussion

## References

### 2024-05-02 index.qmd
added abstract

@@
     email: maturner@stanford.edu
+  - name: Aja Sutton
   - name: Izabela Rezende
@@
     corresponding: true
-bibliography: ../library.bib
+# bibliography: [this.bib, ../library.bib]
+bibliography: this.bib
 classoption: abstract
@@
 abstract: |
-  This is the abstract
+  \noindent Disease transmission leading to pandemics depends on who interacts with whom, but we currently lack a rigorous general understanding of exactly how.
+Metapopulation structure tends to be studied in geospatial terms, with symmetric mixing, i.e., all metapopulation groups have the same symmetric _homophily_, i.e., the same tendency to more or less frequently mix within its own group compared to other groups.
+Heterogeneous mixing, however, has been recognized to matter for epidemiology of HIV/AIDS, the evolution of treatment-resistant malaria,
+  and for the cultural spread of health-protective behaviors such as vaccination, mask wearing, or healthy eating.
+Indeed, social relationships are not just important for predicting disease transmission, but also for predicting the behaviors that may increase or decrease infection risk.
+Therefore, in this paper, we develop an agent-based _coupled contagion_ simulation model of disease and cultural transmission to understand how asymmetric a simple, generic metapopulation structure composed of one minority group and one majority group (by numbers) which have different, asymmetric homophily levels.
+The model is analyzed in both a sexually-transmitted disease-style model where social interactions are structured by an explicit social network, and an airborne disease-style context where mixing is structured only by asymmetric homophily levels.
+We (expect to) find that novel variants with higher transmission probabilities diffuse better when they begin in the minority group than the majority group, and that in this case the minority group will suffer a relatively greater share of the infections as a fraction of the total population.
+Simulated protective measures .
+This model provides a rigorous baseline evaluation of the effect of asymmetric homophily on disease transmission and pandemic preparedness in an idealized setting.
+This baseline provides a backdrop for a rigorous understanding of additional complexity to better evaluate real-world pandemic preparedness, such as psychological/emotional effects like out-group aversion and biased assimilation that can inhibit cultural learning between groups, and such as pathogen evolution in addition to transmission studied here.
 geometry:
@@
 
-## References
+##

### 2024-05-28 index.qmd
Started outlining updated intro for Dourados Reserve

--- 2024-05-02 index.qmd
@@
     email: maturner@stanford.edu
-  - name: Izabela Rezende
-  - name: Benjamin Singer
-  - name: Hannah S. Derwanz
-  - name: James Holland Jones
-    affiliation: Stanford University Doerr School of Sustainability
-    roles: writing
-    corresponding: true
-bibliography: ../library.bib
+bibliography: [this.bib, ../library.bib]
+# bibliography: this.bib
 classoption: abstract
@@
 abstract: |
-  This is the abstract
+  \noindent The COVID-19 pandemic disproportionately affected economically or otherwise marginalized groups, including Indigenous groups.
+In the Dourados Indigenous Reserve in Mato Grosso do Sul, Brazil, it was hypothesized that young men were largely responsible for transmission in and out of the Reserve when they traveled to the neighboring City of Dourados for economic reasons.
+Here we develop an agent-based compartmental model of disease transfer between two groups representing the Reserve and City populations with smaller and larger population-size groups of agents, respectively.
+We consider heterogeneous group-level mixing parameters, where the young Indigenous men have different mobility than other Reserve residents, both of which are different from city resident mobility between Reserve and City.
+We test whether greater mobility among young men could account for observed variant diffusion into the Reserve from the City.
+We also evaluate the potential efficacy of temporary, well-timed economic relief targeted to young men in the Reserve so they would not need to travel the City.
+This work then demonstrates a potential real-world intervention to be used for the next pandemic.
+It also contributes to the theory of epidemics in metapopulations.
+Such scientific predictions are essential for designing a suite of equitable, effective interventions for preventing or mitigating the next pandemic.
 geometry:
@@
 
-Due to rapidly increasing population size and density in a changing climate, the interdisciplinary science of pandemic preparedness is an increasing priority [@Rodo2021; @Adashi2022].
-To complement our biological understanding and preventative tools for mitigating the risk of pandemics, social science is essential to understand how social structure and behavior determines how disease and disease prevention behaviors (such as vaccination[@Amin2017; @Omer2021]) are adopted by individuals in different group-structured populations [@Buckee2021; @Epstein2008; @SmaldinoJones2021].
+To equitably prepare for pandemics in our uncertain world we need to [@Rodo2021; @Adashi2022].
+The global COVID-19 pandemic highlighted the current lack of outcome equality, where a country's inequality (measured by the Gini index) predicted higher burden of COVID-19 infections [@Su2022].
+In this work we contribute to more equitable pandemic preparedness by developing an agent-based compartmental model of COVID-19 transmission between a Brazilian city, Dourados, in Mato Grosso do Sul, and the neighboring Dourados Indigenous Reserve [@DeOliveira2023].
+COVID-19 variant diffusion from Dourados City into the Dourados Reserve was hypothesized to be driven by young Indigenous men traveling between the reserve and the city of Dourados for economic reasons.
+COVID-19 burden was felt disproportionately by Indigenous peoples in Brazil.
+Therefore this case study demonstrates how we can begin to develop scientific predictive models of inter-community disease transmission that could be used to optimize the timing of public health interventions to limit the spread of fitter pathogen variants.
+
+<!-- To complement our biological understanding and preventative tools for mitigating the risk of pandemics, social science is essential to understand how social structure and behavior determines how disease and disease prevention behaviors (such as vaccination[@Amin2017; @Omer2021]) are adopted by individuals in different group-structured populations [@Buckee2021; @Epstein2008; @SmaldinoJones2021].
 Oftentimes in epidemiology, groups are defined by any different group partitions can be relevant to epidemiology, including age (RSV and Polio refs), occupation (frontline/remote refs), and, of course, political party [@VanBavel2020].
@@
 Using this we can infer the effect of different vaccine distribution allotments between the two groups that could possibly better protect one group or the other, or meet other optimal criteria.
-Computational analyses of metapopulation models can be used to create *in silico* experiments that would be impossible to implement in the real world, for instance to predict how different vaccination distribution strategies would best prevent transmission dynamics [@Singer2022].
+Computational analyses of metapopulation models can be used to create in silico experiments that would be impossible to implement in the real world, for instance to predict how different vaccination distribution strategies would best prevent transmission dynamics [@Singer2022].
+-->
 
-Whether a new viral mutation or a cultural adaptation, beneficial *innovations* diffuse differently through different metapopulation structures.
+<!-- Whether a new viral mutation or a cultural adaptation, beneficial innovations diffuse differently through different metapopulation structures.
 Turner and colleagues (2023) showed that novel innovations with higher fitness than an alternative best diffused when first introduced to a model minority group compared to a majority group (in terms of population size), in a metapopulation with a single other majority group [@Turner2023], and diffusion was further optimized when majority groups moderately preferred in-group interactions.
-Preference for in- versus out-group interactions is known as *homophily* [@McPherson2001], and introduced to a simple model metapopulation composed of a minority and majority group (by population size), where there exists a lower-fitness alternative, fitness-biased social learning.
+Preference for in- versus out-group interactions is known as homophily [@McPherson2001], and introduced to a simple model metapopulation composed of a minority and majority group (by population size), where there exists a lower-fitness alternative, fitness-biased social learning.-->
 
-Simple idealized models like we develop here serve as a theoretical ground truth against which we can better understand real-world complexities.
+<!-- Simple idealized models like we develop here serve as a theoretical ground truth against which we can better understand real-world complexities.
 For instance, schistosomaiasis hotspots can be predicted with some accuracy based on data from initial outbreak [@Singer2024], but metapopulation structure may have an underlying effect on the emergence of hotspots that, when accounted for, could increase predictive power of data-driven statistical models.
@@
 For data-driven applications, asymmetric homophily measurements may be easier or more reliable compared to the network structure itself.
+-->
 
-To understand the effect of asymmetric homophily levels on epidemiological dynamics, we developed a computational model where metapopulations are composed of two groups, a minority and majority group defined by population size, where group members interact more or less frequently depending on group membership and group-level homophily levels, and whether interactions are structured by a social network (more like in a sexually-transmitted disease) or by random asymmetric homophilous mixing within and between groups (like in an airborne disease), adapting the models of Turner et al (2023) and Turner, Reynolds, and Jones (2024) for epidemiology by augmenting the social learning dynamic with SIR epidemiological dynamics.
+<!-- To understand the effect of asymmetric homophily levels on epidemiological dynamics, we developed a computational model where metapopulations are composed of two groups, a minority and majority group defined by population size, where group members interact more or less frequently depending on group membership and group-level homophily levels, and whether interactions are structured by a social network (more like in a sexually-transmitted disease) or by random asymmetric homophilous mixing within and between groups (like in an airborne disease), adapting the models of Turner et al (2023) and Turner, Reynolds, and Jones (2024) for epidemiology by augmenting the social learning dynamic with SIR epidemiological dynamics.
+-->
 
-Models like this enable us to rigorously understand how one important mechanism affects the course of disease and optimal intervention strategy.
+<!-- Models like this enable us to rigorously understand how one important mechanism affects the course of disease and optimal intervention strategy.
 The cost of rigor, at this stage, is the exclusion of other important details that can affect disease or cultural transmission, including individual-level general health, access to clean running water, working conditions, and social factors such as asymmetric outgroup aversion levels between groups, which can undermine the adoption of adaptive behaviors [@EhretConstantino2022].
@@
 This work complements a broader effort across evolutionary sciences to understand the interplay between social structure and collective co-adaptation among humans, non-humans, and pathogens [@Bak-Coleman2021; @GalesicEtAl2023].
+-->
 
 ## Model
+
+-   SIR model with
 

### 2024-07-14 index.qmd
updated index and restarted Rproj

--- 2024-05-02 index.qmd
@@
     email: maturner@stanford.edu
-  - name: Izabela Rezende
-  - name: Benjamin Singer
-  - name: Hannah S. Derwanz
-  - name: James Holland Jones
-    affiliation: Stanford University Doerr School of Sustainability
-    roles: writing
-    corresponding: true
-bibliography: ../library.bib
+bibliography: [../library.bib]
+# bibliography: this.bib
 classoption: abstract
@@
 abstract: |
-  This is the abstract
+  \noindent hello
 geometry:
@@
 
-Due to rapidly increasing population size and density in a changing climate, the interdisciplinary science of pandemic preparedness is an increasing priority [@Rodo2021; @Adashi2022].
-To complement our biological understanding and preventative tools for mitigating the risk of pandemics, social science is essential to understand how social structure and behavior determines how disease and disease prevention behaviors (such as vaccination[@Amin2017; @Omer2021]) are adopted by individuals in different group-structured populations [@Buckee2021; @Epstein2008; @SmaldinoJones2021].
-Oftentimes in epidemiology, groups are defined by any different group partitions can be relevant to epidemiology, including age (RSV and Polio refs), occupation (frontline/remote refs), and, of course, political party [@VanBavel2020].
-To understand how social structure affects .
-Using this we can infer the effect of different vaccine distribution allotments between the two groups that could possibly better protect one group or the other, or meet other optimal criteria.
-Computational analyses of metapopulation models can be used to create *in silico* experiments that would be impossible to implement in the real world, for instance to predict how different vaccination distribution strategies would best prevent transmission dynamics [@Singer2022].
+To equitably prepare for pandemics in our uncertain world we need to [@Rodo2021; @Adashi2022].
+The global COVID-19 pandemic highlighted the current lack of outcome equality, where a country's inequality (measured by the Gini index) predicted higher burden of COVID-19 infections [@Su2022].
+Developing computational models can help us understand basic behavioral and social factors underlying the relationship between behavior and disease, but need to be applied to real-world situations because, first and foremost, there are vulnerable populations who need effective mitigation plans for the inevitable next pandemic.
+Second, a generic, yet predictive and application-ready, framework for behavioral-epidemiological metapopulation models can only be developed through making contact with real-world scenarios to hone the model, identifying missing or unnecessary components, and building a repertoire or menu of potential modeling choices that may be important.
+By applying abstract mechanistic models to high-priority real-world problems, we will develop a more broadly useful science of pandemic preparedness.
 
-Whether a new viral mutation or a cultural adaptation, beneficial *innovations* diffuse differently through different metapopulation structures.
-Turner and colleagues (2023) showed that novel innovations with higher fitness than an alternative best diffused when first introduced to a model minority group compared to a majority group (in terms of population size), in a metapopulation with a single other majority group [@Turner2023], and diffusion was further optimized when majority groups moderately preferred in-group interactions.
-Preference for in- versus out-group interactions is known as *homophily* [@McPherson2001], and introduced to a simple model metapopulation composed of a minority and majority group (by population size), where there exists a lower-fitness alternative, fitness-biased social learning.
+In this work we contribute to this program of more equitable pandemic preparedness by developing a compartmental behavioral-epidemiological metapopulation model of COVID-19 transmission between a Brazilian city, Dourados, in Mato Grosso do Sul, and the neighboring Dourados Indigenous Reserve [@DeOliveira2023].
+De Oliveira and co-authors hypothesized that COVID-19 variant diffusion from Dourados City into the Dourados Reserve is driven by young Indigenous men traveling between the reserve and the city of Dourados for economic reasons.
+COVID-19 burden was felt disproportionately by Indigenous peoples in Brazil [@Simionatto2020], but may have been worse in the Dourados Reserve due to its immediate proximity to the Dourados.
+There are many complicating factors that make it difficult to evaluate this hypothesis directly without adding some additional qualifications.
+Here we develop a mechanistic model of epidemiological transmission with asymmetric social mixing that evaluates how best to model group structure in the Reserve.
+In turn, this model is used to evaluate potential economic interventions that target aid to different groups to reduce infections in the reserve.
 
-Simple idealized models like we develop here serve as a theoretical ground truth against which we can better understand real-world complexities.
-For instance, schistosomaiasis hotspots can be predicted with some accuracy based on data from initial outbreak [@Singer2024], but metapopulation structure may have an underlying effect on the emergence of hotspots that, when accounted for, could increase predictive power of data-driven statistical models.
-Metapopulation structure with asymmetric homophily is not yet rigorously understood in an idealized theoretical setting, despite the fact that contact structure can be observed to affect transmission dynamics of sexually transmitted disease [@Morris1991], the spread of antibiotic-resistant bacteria across hospital wards [@Shapiro2020], the emergence and diffusion of drug-resistant malaria [@Ariey2003], the spread of schistosomiasis [@Ciddio2017], and provide disease reservoirs that drove the spread of tuberculosis and COVID-19 between some prison and community populations [@Sacchi2015; @Ryckman2021].
-Contact structure has been rigorously shown to shape epidemiological dynamics [@Arthur2017; @Großmann2021], but not in settings where social networks are defined by asymmetric homophily, specifically.
-For data-driven applications, asymmetric homophily measurements may be easier or more reliable compared to the network structure itself.
+-   De Oliveira and colleagues tracked the spread of different SARS-CoV-2 variants by village and by variant.
 
-To understand the effect of asymmetric homophily levels on epidemiological dynamics, we developed a computational model where metapopulations are composed of two groups, a minority and majority group defined by population size, where group members interact more or less frequently depending on group membership and group-level homophily levels, and whether interactions are structured by a social network (more like in a sexually-transmitted disease) or by random asymmetric homophilous mixing within and between groups (like in an airborne disease), adapting the models of Turner et al (2023) and Turner, Reynolds, and Jones (2024) for epidemiology by augmenting the social learning dynamic with SIR epidemiological dynamics.
+![Time series of COVID-19 spread through the Dourados Indigenous Reserve, with significant events highlighted.
+The first B.1.1 variant cases were introduced by the Guarani ethnicity (1), followed about one month later by a rapid rise in cases in Jaguapiru among the Terena ethnicity(2); several months later a rapid rise in Bororó was caused by a rapid rise among Guarani (3).
+Zeta was introduced at the end of 2020 (4), with a rapid rise around the holiday season especially in Jaguapiru among Guarani (5); later in February cases rose in Bororó among the Terena (6).](Figures/Time%20Series.pdf){width=80%}
 
-Models like this enable us to rigorously understand how one important mechanism affects the course of disease and optimal intervention strategy.
-The cost of rigor, at this stage, is the exclusion of other important details that can affect disease or cultural transmission, including individual-level general health, access to clean running water, working conditions, and social factors such as asymmetric outgroup aversion levels between groups, which can undermine the adoption of adaptive behaviors [@EhretConstantino2022].
-However, this initial mechanistic model ideally sets in motion a series of follow-on studies to better understand which of these additional factors alter model predictions to what extent.
-Model refinements and specializations can be added for specific epidemiological contexts, different scenarios and intervention strategies can be tested computationally for optimal policies given available information.
-More shortcomings, refinements, and specializations can follow that, with the process repeating as necessary.
-This work complements a broader effort across evolutionary sciences to understand the interplay between social structure and collective co-adaptation among humans, non-humans, and pathogens [@Bak-Coleman2021; @GalesicEtAl2023].
+-   The goal of our model is two-fold. First, to identify parameters that could re-create the time series plots in Figure 1.
+-   Second, we aim to use a parameterized model to understand how different intervention strategies could reduce infections among Reserve residents.
+For example, perhaps since Terena are suspected to travel to the
+-   While we know a lot about transmission in the reserve, and the work of de Oliveira and colleagues advanced our understanding of transmission in the Reserve, we do not have reliable estimates of exactly how many cases there were per ethnicity and village.
+We also have only word-of-mouth information about the frequency of travel to the city for each ethnicity and location.
+Our model motivates the need for a more thorough understanding of behavior and epidemiological parameters in the Dourados Indigenous Reserve, and points the way towards future research in other Indigenous communities, and more complex situations.
 
 ## Model
+
+![We can model the relevant groups for mixing as either village residence or ethnicity.
+Transmission, mixing, and mobility parameters would be set accordingly.
+The $C_g$ parameters would be reduced if economic aid were allocated to keep group $g$ on the Reserve during outbreaks.](Figures/ModelSketch.pdf){fig-align="center"}
 
@@
 ## Discussion
-
-## References

### 2024-07-18 ms.tex
updated equations with mixing, city travel, and reinfection terms

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  abstract]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.1in,letterpaper]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[sorting=none, style=numeric-comp]{biblatex}
\addbibresource{../library.bib}
\usepackage{bookmark}

\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Predicting the effect of candidate interventions for mitigating disease spread in asymmetric mixing metapopulations},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Predicting the effect of candidate interventions for mitigating
disease spread in asymmetric mixing metapopulations}
\author{Matthew Adam Turner}
\date{}

\begin{document}
\maketitle
\begin{abstract}
\noindent hello
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{introduction}

To equitably prepare for pandemics in our uncertain world we need to
\autocite{Rodo2021,Adashi2022}. The global COVID-19 pandemic highlighted
the current lack of outcome equality, where a country's inequality
(measured by the Gini index) predicted higher burden of COVID-19
infections \autocite{Su2022}. Here we contribute to this program of more
equitable pandemic preparedness by developing a compartmental
behavioral-epidemiological metapopulation model of COVID-19 transmission
between a Brazilian city, Dourados, in Mato Grosso do Sul, and the
neighboring Dourados Indigenous Reserve \autocite{DeOliveira2023}.
COVID-19 burden was felt disproportionately by Indigenous peoples in
Brazil \autocite{Simionatto2020}, but may have been worse in the
Dourados Reserve due to its immediate proximity to the Dourados. Nearly
all residents of the reserve live in domiciles in one of two villages in
the Reserve, Jaguapiru and Bororó. Jaguapiru is closer overall to the
city of Dourados, sharing a much longer border with the city limits than
Bororó (Figure 1). Indigenous peoples on the reserve are predominantly
one of two ethnicities, Guarani and Terena. Anecdotally, Terena are
known for being better educated with somewhat more robust physicality
and, perhaps, better resistance to disease. They may travel to the city
more frequently for trade and other purposes.

De Oliveira and co-authors hypothesized that COVID-19 variant diffusion
from Dourados City into the Dourados Reserve is driven by young
Indigenous men traveling between the reserve and the city of Dourados
for economic reasons. This is a reasonable hypothesis given available
data.

\begin{figure}[H]

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{Figures/ReserveMap.pdf}

}

\caption{Map of Dourados Indigenous Reserve villages (Bororó in light
blue, Jaguapiru in cyan) and the city of Dourados (dark blue). Note
Jaguapiru shares a much longer boundary with the reserve than Dourados,
which facilitates mobility in and out of the reserve for both Reserve
and city residents. Maps combined from Schnaufer et al.~(2023) and
Cavalcante (2019)}

\end{figure}%%
\begin{figure}[H]

{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{Figures/Time Series.pdf}

}

\caption{Time series of COVID-19 spread through the Dourados Indigenous
Reserve, with significant events highlighted. The first B.1.1 variant
cases were introduced by the Guarani ethnicity (1), followed about one
month later by a rapid rise in cases in Jaguapiru among the Terena
ethnicity(2); several months later a rapid rise in Bororó was caused by
a rapid rise among Guarani (3). Zeta was introduced at the end of 2020
(4), with a rapid rise around the holiday season especially in Jaguapiru
among Guarani (5); later in February cases rose in Bororó among the
Terena (6).}

\end{figure}%

Despite some evidence to support the hypothesis that young male
Indigenous mobility can explain transmission into the reserve, there are
many complicating factors that make it difficult to evaluate this
hypothesis directly without adding some additional qualifications. There
may be other social factors that may be necessary for the time series
observations of De Oliveira and co-authors. Furthermore, there may be
less significant factors driving transmission in this model that may be
of greater importance in future outbreaks of different diseases than
SARS-CoV-2 in the reserve. Therefore, here we expand the scope of the
hypothesis to include hypotheses about how social groups, namely village
residency and Indigenous ethnic group, affect transmission in the
reserve and explain observed outbreaks of variants B.1.1 and Zeta in the
Reserve. To evaluate the hypotheses, presented in the next paragraph, we
develop a mechanistic model of epidemiological transmission with
asymmetric social mixing that evaluates how best to model group
structure in the Reserve. In turn, this model is used to evaluate
potential economic interventions that target aid to different groups to
reduce infections in the reserve.

\begin{itemize}
\item
  De Oliveira and colleagues tracked the spread of different SARS-CoV-2
  variants by village and by variant.
\item
  The goal of our model is two-fold. First, to identify parameters that
  could re-create the time series plots in Figure 1.
\item
  Second, we aim to use a parameterized model to understand how
  different intervention strategies could reduce infections among
  Reserve residents. For example, perhaps since Terena are suspected to
  travel to the
\item
  While we know a lot about transmission in the reserve, and the work of
  de Oliveira and colleagues advanced our understanding of transmission
  in the Reserve, we do not have reliable estimates of exactly how many
  cases there were per ethnicity and village. We also have only
  word-of-mouth information about the frequency of travel to the city
  for each ethnicity and location. Our model motivates the need for a
  more thorough understanding of behavior and epidemiological parameters
  in the Dourados Indigenous Reserve, and points the way towards future
  research in other Indigenous communities, and more complex situations.
  Here we only consider Indigenous mobility, but many people from the
  city travel to the reserve for commerce as well.
\end{itemize}

\section{Model}\label{model}

\begin{figure}[H]

{\centering \includegraphics{Figures/ModelSketch.pdf}

}

\caption{We can model the relevant groups for mixing as either village
residence or ethnicity. Transmission, mixing, and mobility parameters
would be set accordingly. The \(C_g\) parameters would be reduced if
economic aid were allocated to keep group \(g\) on the Reserve during
outbreaks.}

\end{figure}%

\subsection{Equations}\label{equations}

\begin{align*}
  \frac{dS_g}{dt} &= -\left(\tau c_{gg'} I_g + c_{g'g}I_{g'} + \gamma_g \right)S_g + r R_g   \\
  \frac{dI_g}{dt} &= \left(\tau c_{gg'} I_g + c_{g'g}I_{g'} + \gamma_g \right)S_g - \rho I_g \\
  \frac{dR_g}{dt} &= \rho I_g - r R_g
    \numberthis \label{eq:DiseaseContagion}
\end{align*}
\noindent

\section{Analysis (Sketch)}\label{analysis-sketch}

\begin{figure}[H]

{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch.pdf}

}

\caption{Global reduction in travel to and from the city could have
differential effects on reduction in incidence for different groups.}

\end{figure}%

\section{Discussion}\label{discussion}

\printbibliography[title=References]

\end{document}

### 2024-07-18 ms.tex
added equations and started disc outline

@@
 \usepackage{bookmark}
-
+\usepackage{xcolor}
+\definecolor{myorange}{RGB}{240, 96, 0}
+\newcommand{\mt}[1]{{\textcolor{myorange} {({\tiny MT:} #1)}}}
 \newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
@@
 
-\caption{We can model the relevant groups for mixing as either village
+\caption{\mt{UPDATE THIS CAPTION AND THE FIGURE TO MATCH EQUATIONS. ADD PARAMETERS TABLE.}
+  We can model the relevant groups for mixing as either village
 residence or ethnicity. Transmission, mixing, and mobility parameters
@@
 
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch.pdf}
+{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ReductionByGroup.pdf}
 
@@
 
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByStrategy.pdf}
+
+}
+
+\caption{We hypothesize that overall reduction in disease incidence ($\Delta I$, y-axis)  depends on 
+strategy to distribute the total aid to reduce travel between Dourados Reserve and city ($\gamma$, x-axis).}
+
+\end{figure}%
+
 \section{Discussion}\label{discussion}
+
+\begin{itemize}
+  \item 
+    Application to other problematic diseases in the Reserve, e.g., TB, STDs,\ldots 
+  \item
+    The data is not necessarily representative and we had to use educated guesses
+    for mixing parameters.
+\end{itemize}
 

### 2024-07-22 ms.tex
wrote formalism paragraph, on to sketch of computational experiment

--- 2024-07-18 ms.tex
@@
 \makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
+% \@ifundefined{KOMAClassName}{% if non-KOMA class
+  % \IfFileExists{parskip.sty}{%
+  %   \usepackage{parskip}
+  % }{% else
+  %   % \setlength{\parindent}{pt}
+  %   \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+% }{% if KOMA class
+  % \KOMAoptions{parskip=half}}
 \makeatother
@@
 \title{Predicting the effect of candidate interventions for mitigating
-disease spread in asymmetric mixing metapopulations}
+disease spread in asymmetric-mixing metapopulations}
 \author{Matthew Adam Turner}
-\date{}
+\date{\today}
 
@@
 
-\begin{figure}[H]
+\mt{IF TERENA ARE MINORITY, HAVE HIGH MOBILITY TO CITY, AND $c_{GT} > c_{TG}$
+THERE COULD BE A MINORITY INCUBATOR-TYPE EFFECT~\autocite{Turner2023}}
 
-{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{Figures/ReserveMap.pdf}
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:Map} ABOUT HERE ]} \\
+\end{center}
 
-}
-
-\caption{Map of Dourados Indigenous Reserve villages (Bororó in light
-blue, Jaguapiru in cyan) and the city of Dourados (dark blue). Note
-Jaguapiru shares a much longer boundary with the reserve than Dourados,
-which facilitates mobility in and out of the reserve for both Reserve
-and city residents. Maps combined from Schnaufer et al.~(2023) and
-Cavalcante (2019)}
-
-\end{figure}%%
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{Figures/Time Series.pdf}
-
-}
-
-\caption{Time series of COVID-19 spread through the Dourados Indigenous
-Reserve, with significant events highlighted. The first B.1.1 variant
-cases were introduced by the Guarani ethnicity (1), followed about one
-month later by a rapid rise in cases in Jaguapiru among the Terena
-ethnicity(2); several months later a rapid rise in Bororó was caused by
-a rapid rise among Guarani (3). Zeta was introduced at the end of 2020
-(4), with a rapid rise around the holiday season especially in Jaguapiru
-among Guarani (5); later in February cases rose in Bororó among the
-Terena (6).}
-
-\end{figure}%
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:DeOliveiraSeries} ABOUT HERE ]} \\
+\end{center}
 
@@
 
-\section{Model}\label{model}
+\section{Model}\label{sec:model}
+
+To understand and predict how different candidate economic interventions could
+reduce transmission in the Dourados Reserve, we develop a mechanistic model of
+transmission in the Reserve implemented as a set of compartmental
+susceptible-infected-recovered system (SIR) of differential equations.  There
+are six compartments in the differential equations, one $S_g$, $I_g$, and $R_g$ for
+each group, $g$.  We evaluate different interventions by varying the reduction in
+Indigenous travel between the Dourados reserve and the city, which in the real
+world would be achieved by economic aid that reduces the need for Indigenous
+travel to the city. First we test the effect of global reduction in travel on
+infections in the Reserve, where aid is distributed equally across all possible
+groupings. Second, we test different strategies for distributing aid to
+different groups to understand which, if any, are more effective than equal aid
+distribution. Finally, we test infectin prevention under both global and
+group-specific intervention strategies for a stylized version of the Zeta
+variant that allows probabilistically for re-infection. Although the model is
+focused on COVID-19 transmission in the Dourados Indigenous Reserve, the
+model is abstract. What we learn in our analyses could therefore transfer to
+other diseases in the Reserve or other socio-epidemiological contexts.
+
+\subsection{Model dynamics and formalism}
+\label{sub:modelDynamics}
+
+Our model represents interactions within and between Dourados Indigenous 
+Reserve groups, and with members of the city of Dourados 
+(Figure~\ref{fig:ModelSketch}). To represent
+interactions between groups we define two possible groupings, by village
+(Figure~\ref{fig:ModelSketch}, Model 1) and by ethnicity
+(Figure~\ref{fig:ModelSketch}, Model 2). The within-group contact rate 
+is $c_{gg}$ for focal group $g$. The contact rate between the focal group and
+its outgroup, $g'$, is $c_{gg'}$. If in Model 2 the focal group is Terena, then
+the within-group contact rate is $c_{TT}$ and the between-group contact rate,
+representing how frequently Terena interact with Guarani is $c_{TG}$, Each of these groupings can be further
+subdivided by the alternative Model grouping, represented in 
+Figure~\ref{fig:ModelSketch} by the inset pie chart diagrams in each group's
+circle. Reserve resident mobility is represented by a contact-infection rate,
+$\gamma_g$, which represents how frequently a susceptible member of group $g$
+brings an infection into the Reserve. The exact compartments, $S$, $I$, and $R$
+for the city population are ignored.
+
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:ModelSketch} ABOUT HERE ]} \\
+\end{center}
+
+This contact structure, combined with a global transmission rate $\tau$,
+recovery rate, $\rho$, and re-susceptibility rate, $r$, results in a system of
+six differential equations for each of the six compartments
+(Equation~\ref{eq:DiffEqs}). $\tau$ and $\rho$ will be set to reasonable values
+for main analyses, but varied for sensitivity analyses.  \mt{Q: What constitutes
+a ``reasonable value'' here? Values Aja used were for ABM---what about for a
+diffeq model like this?}. To model increased re-infections observed with the
+Zeta variant, we also include a term $r R_g$ that is positive in the
+$\frac{dS_g}{dt}$ equation and negative in the $\frac{dR_g}{dt}$ term to
+transfer recovered individuals from the recovered compartment for group $g$ to
+the susceptible compartment for that group. If $r=0$, then there will be no
+re-infections. Otherwise there may be re-infections, which could have complex
+interactions with interaction structure.
+
+\begin{align*}
+  \frac{dS_g}{dt} &= -m_{gg'} \tau S_g + r R_g   \\
+  \frac{dI_g}{dt} &=  m_{gg'} \tau S_g - \rho I_g \\
+  \frac{dR_g}{dt} &= \rho I_g - r R_g,\quad\text{ where } \\[1em]
+          m_{gg'} &= \left(c_{gg'} I_g + c_{g'g}I_{g'} + \gamma_g \right)
+    \numberthis \label{eq:DiffEqs}
+\end{align*}
+\noindent
+
+\vspace{0.5em}
+\begin{center}
+{[ TABLE \ref{tab:Variables} ABOUT HERE ]} \\
+\end{center}
+
+\subsection{Computational analysis}
+
+\section{Analysis (Sketch)}\label{analysis-sketch}
+
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:GlobalTravelReduction} ABOUT HERE ]} \\
+\end{center}
+
+\vspace{0.5em}
+\begin{center}
+{[ FIGURE \ref{fig:GroupSpecificTravelReduction} ABOUT HERE ]} \\
+\end{center}
+
+\section{Discussion}\label{discussion}
+
+\begin{itemize}
+  \item 
+    Application to other problematic diseases in the Reserve, e.g., TB, STDs,\ldots 
+  \item
+    The data is not necessarily representative and we had to use educated guesses
+    for mixing parameters.
+  \item
+    Have not yet considered more complex epidemiological and behavioral
+    dynamics. For example, group structure can affect the coupled contagion of
+    disease and behaviors that could encourage or impede the spread of
+    disease~\autocite{SmaldinoJones2021}. Our compartmental model
+    incorporated group structure and contact rate heterogeneity, but finer
+    details of the social network structure can induce complex
+    epidemiological dynamics~\autocite{Arthur2017}. In the case of Dourados
+    and possibly other Indenous communities is the tendency for several
+    generations of families to live in close proximinty to one another,
+    often communally with other families as well, making social isolation
+    culturally unpalatable and, consequentially, unfeasible. Adaptation to
+    evolving, uncertain epidemiological social and environmental contexts is
+    one further factor that can complicate epidemiological
+    dynamics~\autocite{Arthur2021} and willingness to learn beneficial public
+    health behaviors from others~\autocite{Turner2023a}.
+\end{itemize}
+
+\printbibliography[title=References]
+
+\clearpage
+
+\section{Tables}
+
+\begin{table}[h]
+  \centering
+  \caption{\textbf{Model variables, description and values taken.} The
+  compartmental model described in Equation~\ref{eq:DiffEqs} includes several
+social mixing, behavioral, and epidemiological variables described here.}
+  \label{tab:Variables}
+  \begin{tabular}{rlp{1.3in}} 
+    \toprule
+    variable & description & values \\ 
+    \midrule 
+    $S_g,I_g,R_g$ & Susceptible, infected, recovered compartments by group & $[0.0,1.0]$ \\
+    $c_{gg'},c_{g'g}$ & Asymmetric inter-group contact rates & $[0.0,1.0]$ \\
+    $\gamma_g$ & Group-level incidence-contact rate with city residents &
+      $\{0.0,0.1,\ldots,0.4\}$ \\
+    $\rho$ & Global recovery rate & $[0.0,1.0]$ \\
+    $\tau$ & Global ransmission rate & $[0.0,1.0]$ \\
+    $r$    & Re-susceptible probability, non-zero only for Zeta & $0.0,0.5$ \\
+    $g,g'$ & Reference group ($g$) and out-group $g'$ & Terena, Guarani; Bororó, Jaguapiru \\
+    \bottomrule
+  \end{tabular} 
+\end{table}
+
+\section{Figures}
 
@@
 
+\centering
+\caption{\textbf{Map of Dourados Indigenous Reserve villages.} Bororó is light
+blue, Jaguapiru cyan, and the city of Dourados dark blue. Note
+Jaguapiru shares a much longer boundary with the reserve than Dourados,
+which facilitates mobility in and out of the reserve for both Reserve
+and city residents. Maps combined from Cavalcante (2019) for the underlying
+topographic map and Schnaufer et al.~(2023) for the villages, city, and 
+sample collection points (clinics) layer.}
+\label{fig:Map}
+\vspace{0.3em}
+{\includegraphics[width=0.8\textwidth,height=\textheight]{Figures/ReserveMap.pdf}}
+\end{figure}%%
+
+\begin{figure}[H]
+\caption{Time series of COVID-19 spread through the Dourados Indigenous
+Reserve, with significant events highlighted. The first B.1.1 variant
+cases were introduced by the Guarani ethnicity (1), followed about one
+month later by a rapid rise in cases in Jaguapiru among the Terena
+ethnicity (2); several months later a rapid rise in Bororó was caused by
+a rapid rise among Guarani (3). Zeta was introduced at the end of 2020
+(4), with a rapid rise around the holiday season especially in Jaguapiru
+among Guarani (5); later in February cases rose in Bororó among the
+Terena (6).}
+\label{fig:DeOliveiraSeries}
+\vspace{0.3em}
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{Figures/Time Series.pdf}
+
+}
+\end{figure}%
+
+\begin{figure}[H]
+
+\caption{
+  We can model the relevant groups for mixing as either village
+residence or ethnicity. Mixing, and mobility parameters
+would be set accordingly. Mixing is asymmetric, with $c_{gg}$ representing the
+contact rate within group $g$, and each $c_{gg'}$ representing
+the contact rate at which a member of group $g$ encounters a member of the
+outgroup $g'$, e.g., $c_{GT}$ is the contact rate at which Guarani encounter
+members of the Terena. The $\gamma_g$ parameters would be reduced through
+economic aid to group $g$ for them to remain on the Reserve during
+outbreaks. \mt{Looks like some labels are cut off; will fix in later draft.}}
+\label{fig:ModelSketch}
+\vspace{0.3em}
 {\centering \includegraphics{Figures/ModelSketch.pdf}
@@
 }
+\end{figure}%
 
-\caption{We can model the relevant groups for mixing as either village
-residence or ethnicity. Transmission, mixing, and mobility parameters
-would be set accordingly. The \(C_g\) parameters would be reduced if
-economic aid were allocated to keep group \(g\) on the Reserve during
-outbreaks.}
+\begin{figure}[H]
+
+\caption{Global reduction in travel to and from the city could have
+Infection reduction ($\Delta I$, y-axis) for different groups could be
+different as global reduction in travel to and from the city decreases
+($\Delta \gamma$, x-axis).}
+\label{fig:GlobalTravelReduction}
+\vspace{0.3em}
+{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ReductionByGroup.pdf}
+
+}
 
@@
 
-\subsection{Equations}\label{equations}
-
-\begin{align*}
-  \frac{dS_g}{dt} &= -\left(\tau c_{gg'} I_g + c_{g'g}I_{g'} + \gamma_g \right)S_g + r R_g   \\
-  \frac{dI_g}{dt} &= \left(\tau c_{gg'} I_g + c_{g'g}I_{g'} + \gamma_g \right)S_g - \rho I_g \\
-  \frac{dR_g}{dt} &= \rho I_g - r R_g
-    \numberthis \label{eq:DiseaseContagion}
-\end{align*}
-\noindent
-
-\section{Analysis (Sketch)}\label{analysis-sketch}
-
 \begin{figure}[H]
 
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch.pdf}
+\caption{We hypothesize that overall reduction in disease incidence ($\Delta I$, y-axis)  depends on 
+strategy to distribute the total aid to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
+\label{fig:GroupSpecificTravelReduction}
+\vspace{0.3em}
+{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByStrategy.pdf}
 
@@
 
-\caption{Global reduction in travel to and from the city could have
-differential effects on reduction in incidence for different groups.}
-
 \end{figure}%
 
-\section{Discussion}\label{discussion}
-
-\printbibliography[title=References]
-
 \end{document}

### 2024-07-24 ms.tex
nearing complete model desc & proj plan

@@
 
+We evaluate hypothetical interventions to limit the spread of
+infection in the Reserve by simulating these interventions using the formal
+system explained above. Our main outcome variable is the reduction in simulated
+COVID-19 incidence among Indigenous peoples compared to the no-intervention
+case, defined 
+\begin{equation}
+  \Delta I = I(\text{no intervention}) - I(\text{intervention } i).
+\end{equation}
+\noindent
+A preliminary step is to find parameterizations that reproduce the observed
+outbreaks of B.1.1 and Zeta within the Reserve to obtain a baseline estimate of
+infection size $I(\text{no intervention})$. 
+To this end we hypothesize that re-infections could have driven the observed
+greater spread of the Zeta strain, though it could also be due to the holiday
+season around the end of 2020 and beginning of 2021.
+
+Interventions center around reducing the mobility of Indigenous people between
+Dourados city and Reserve; the total change in mobility, denoted $\Delta
+\gamma$, is the total reduction of the parameter $\gamma_g$ across all 
+groups, i.e., 
+\begin{equation}
+  \Delta \gamma = \sum_g \Delta \gamma_g.
+\end{equation}
+\noindent
+Different interventions allocate $\Delta \gamma$ to different degrees. 
+One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
+equal. Another named strategy is the \emph{80-20} strategy, where the
+reduction in mobility in group $g$ is 80\% of the total, and reduction in
+group $g'$ mobility is 20\% of the total, i.e., $\Delta\gamma_g / \Delta\gamma =
+0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$ To create alternative
+strategies, both the total reduction and the ratio of group-level reduction to 
+total reduction can be changed. We finally evaluate the hypothesis that
+interventions that may be effective for strains with no re-infection (like
+B.1.1) will not be equally effective for strains with re-infection (like Zeta),
+and vice-versa. This will highlight the importance of accounting for social and
+epidemiological context, alike, in pandemic preparedness via this program of
+computational evaluation of intervention strategies.
+
 \section{Analysis (Sketch)}\label{analysis-sketch}

### 2024-07-24 ms.tex
connecting computational analysis subsec with analysis sketch

@@
 
-To equitably prepare for pandemics in our uncertain world we need to
+To equitably prepare for pandemics in our uncertain world we
+need to predict the effect of candidate interventions to mitigate novel spillover
+disease and novel variants of existing disease
 \autocite{Rodo2021,Adashi2022}. The global COVID-19 pandemic highlighted
@@
 equitable pandemic preparedness by developing a compartmental
-behavioral-epidemiological metapopulation model of COVID-19 transmission
-between a Brazilian city, Dourados, in Mato Grosso do Sul, and the
+behavioral-epidemiological metapopulation model with an immediate application:
+disease transmission between a Brazilian city, Dourados, in the state of 
+Mato Grosso do Sul, and the
 neighboring Dourados Indigenous Reserve \autocite{DeOliveira2023}.
-COVID-19 burden was felt disproportionately by Indigenous peoples in
-Brazil \autocite{Simionatto2020}, but may have been worse in the
-Dourados Reserve due to its immediate proximity to the Dourados. Nearly
+COVID-19 burden was felt disproportionately by Indigenous peoples throughout
+Brazil \autocite{Simionatto2020}, but the burden may have been even more 
+inequitable in the
+Dourados Reserve due to its immediate proximity to the city of Dourados. Nearly
 all residents of the reserve live in domiciles in one of two villages in
@@
 city of Dourados, sharing a much longer border with the city limits than
-Bororó (Figure 1). Indigenous peoples on the reserve are predominantly
-one of two ethnicities, Guarani and Terena. Anecdotally, Terena are
-known for being better educated with somewhat more robust physicality
-and, perhaps, better resistance to disease. They may travel to the city
-more frequently for trade and other purposes.
-
+Bororó (Figure \ref{fig:Map}). Indigenous peoples on the reserve are predominantly
+one of two ethnicities, Guarani and Terena. 
+% Anecdotally, Terena are
+% known for being better educated with somewhat more robust physicality
+% and, perhaps, better resistance to disease. 
+% Anecdotally Terena may travel to the city more frequently for trade and other purposes,
+% however De Olivera and colleagues 
 De Oliveira and co-authors hypothesized that COVID-19 variant diffusion
@@
 Indigenous men traveling between the reserve and the city of Dourados
-for economic reasons. This is a reasonable hypothesis given available
-data.
+for economic reasons. This is a reasonable starting hypothesis given available
+data, but it points to potentially richer, more explanatory mechanisms and 
+a more detailed set of hypotheses. Understanding transmission in the Reserve, and
+evaluating possible interventions based on this understanding, could mitigate
+the next pandemic that affects the Dourados Reserve, and lessons learned here could
+transfer to pandemic preparedness planning in other contexts.
 
-\mt{IF TERENA ARE MINORITY, HAVE HIGH MOBILITY TO CITY, AND $c_{GT} > c_{TG}$
-THERE COULD BE A MINORITY INCUBATOR-TYPE EFFECT~\autocite{Turner2023}}
+Despite some evidence to support the hypothesis that young male
+Indigenous mobility can explain transmission into the reserve, there are
+many complicating factors that make it difficult to evaluate this
+hypothesis directly without adding some additional qualifications. 
+There
+may be other social factors that may be necessary for the time series
+observations of De Oliveira and co-authors. 
+Furthermore, there may be
+less significant factors driving transmission in this model that may be
+of greater importance in future outbreaks of different diseases than
+SARS-CoV-2 in the reserve. 
+Therefore, here we expand the scope of the
+hypothesis to include hypotheses about how social groups, namely village
+residency and Indigenous ethnic group, affect transmission in the
+reserve and explain observed outbreaks of variants B.1.1 and Zeta in the
+Reserve. To evaluate the hypotheses, presented in the next paragraph, we
+develop a mechanistic model of epidemiological transmission with
+asymmetric social mixing that evaluates how best to model group
+structure in the Reserve. In turn, this model is used to evaluate
+potential economic interventions that target aid to different groups to
+reduce infections in the reserve.
 
@@
 \end{center}
-
-Despite some evidence to support the hypothesis that young male
-Indigenous mobility can explain transmission into the reserve, there are
-many complicating factors that make it difficult to evaluate this
-hypothesis directly without adding some additional qualifications. There
-may be other social factors that may be necessary for the time series
-observations of De Oliveira and co-authors. Furthermore, there may be
-less significant factors driving transmission in this model that may be
-of greater importance in future outbreaks of different diseases than
-SARS-CoV-2 in the reserve. Therefore, here we expand the scope of the
-hypothesis to include hypotheses about how social groups, namely village
-residency and Indigenous ethnic group, affect transmission in the
-reserve and explain observed outbreaks of variants B.1.1 and Zeta in the
-Reserve. To evaluate the hypotheses, presented in the next paragraph, we
-develop a mechanistic model of epidemiological transmission with
-asymmetric social mixing that evaluates how best to model group
-structure in the Reserve. In turn, this model is used to evaluate
-potential economic interventions that target aid to different groups to
-reduce infections in the reserve.
 
@@
 group $g'$ mobility is 20\% of the total, i.e., $\Delta\gamma_g / \Delta\gamma =
-0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$ To create alternative
+0.8$ and $\Delta\gamma_{g'} / \Delta\gamma = 0.2$. To create alternative
 strategies, both the total reduction and the ratio of group-level reduction to 
@@
 \begin{center}
-{[ FIGURE \ref{fig:GlobalTravelReduction} ABOUT HERE ]} \\
+{[ FIGURE \ref{fig:GroupSpecificTravelReduction} ABOUT HERE ]} \\
 \end{center}
@@
 \begin{center}
-{[ FIGURE \ref{fig:GroupSpecificTravelReduction} ABOUT HERE ]} \\
+{[ FIGURE \ref{fig:GlobalTravelReduction} ABOUT HERE ]} \\
 \end{center}

### 2024-07-25 ms.tex
variables and cpu analysis updates

@@
 \begin{center}
-{[ TABLE \ref{tab:Variables} ABOUT HERE ]} \\
+{[ TABLE \ref{tab:ModelVariables} ABOUT HERE ]} \\
 \end{center}
@@
   \Delta I = I(\text{no intervention}) - I(\text{intervention } i).
+  \label{eq:DeltaI}
 \end{equation}
@@
   \Delta \gamma = \sum_g \Delta \gamma_g.
+  \label{eq:DeltaGamma}
 \end{equation}
 \noindent
-Different interventions allocate $\Delta \gamma$ to different degrees. 
+Different interventions allocate $\Delta \gamma$ (Equation~\ref{eq:DeltaGamma})
+in different ways to hypothetically result in different case reductions, $\Delta
+I$ (Equation~\ref{eq:DeltaI}). 
 One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
@@
 strategies, both the total reduction and the ratio of group-level reduction to 
-total reduction can be changed. We finally evaluate the hypothesis that
-interventions that may be effective for strains with no re-infection (like
-B.1.1) will not be equally effective for strains with re-infection (like Zeta),
-and vice-versa. This will highlight the importance of accounting for social and
-epidemiological context, alike, in pandemic preparedness via this program of
-computational evaluation of intervention strategies.
+total reduction can be changed. In addition to total case reduction, we 
+measure relative group-level case reduction, $\Delta I_g$, 
+for each considered strategy to understand whether case reduction is equitable
+between villages and ethnicities under different strategies. 
+We finally evaluate the hypothesis that interventions that may be effective for
+strains with no re-infection (like B.1.1) will not be equally effective for
+strains with re-infection (like Zeta), and vice-versa. 
+% This will highlight the importance of accounting for social and
+% epidemiological context, alike, in predicting the effectiveness of mitigation
+% strategies for pandemic preparedness.
 
@@
 social mixing, behavioral, and epidemiological variables described here.}
-  \label{tab:Variables}
+  \label{tab:ModelVariables}
   \begin{tabular}{rlp{1.3in}} 

### 2024-07-25 ms.tex
harmonized figures with analysis variables; added analysis variable table

@@
 infection in the Reserve by simulating these interventions using the formal
-system explained above. Our main outcome variable is the reduction in simulated
-COVID-19 incidence among Indigenous peoples compared to the no-intervention
-case, defined 
+system explained above. Our main outcome variable is the mean reduction 
+in simulated COVID-19 incidence among group $g$ as a fraction of the no-intervention
+cumulative cases, defined at the group level as
 \begin{equation}
-  \Delta I = I(\text{no intervention}) - I(\text{intervention } i).
+  \langle \Delta I_g \rangle = \left\langle \frac{I^\mathrm{Tot}_g(\text{no intervention}) -
+    I^\mathrm{Tot}_g(\text{intervention } i)}{I^\mathrm{Tot}_g(\text{no
+  intervention})} \right\rangle.
+  \label{eq:DeltaI_byGroup}
+\end{equation}
+\noindent
+The mean total expected reduction of cumulative cases is 
+\begin{equation}
+  \langle \Delta I^\mathrm{Tot} \rangle = \left \langle \sum_g \Delta I_g \right \rangle = \sum_g \left \langle \Delta I_g \right \rangle 
   \label{eq:DeltaI}
@@
 \noindent
-A preliminary step is to find parameterizations that reproduce the observed
-outbreaks of B.1.1 and Zeta within the Reserve to obtain a baseline estimate of
-infection size $I(\text{no intervention})$. 
-To this end we hypothesize that re-infections could have driven the observed
-greater spread of the Zeta strain, though it could also be due to the holiday
-season around the end of 2020 and beginning of 2021.
+by linearity of expectations.
+In the case of deterministic differential equations, these expectations are just the
+values themselves.  We can define A preliminary step is to find parameterizations
+that reproduce the observed outbreaks of B.1.1 and Zeta within the Reserve to obtain
+a baseline estimate of infection size $I(\text{no intervention})$.  To this end we
+hypothesize that re-infections could have driven the observed greater spread of the
+Zeta strain, though it could also be due to the holiday season around the end of
+2020 and beginning of 2021.
 
@@
 Different interventions allocate $\Delta \gamma$ (Equation~\ref{eq:DeltaGamma})
-in different ways to hypothetically result in different case reductions, $\Delta
-I$ (Equation~\ref{eq:DeltaI}). 
+in different ways to hypothetically result in different case reductions for each
+group and overall (Equation~\ref{eq:DeltaI_byGroup} and Equation~\ref{eq:DeltaI},
+respectfully; see Table~\ref{tab:AnalysisVariables} for a summary of these analysis
+variables).
 One strategy we test is the \emph{Equal} strategy, where all $\gamma_g$ are
@@
 \begin{center}
-{[ FIGURE \ref{fig:GroupSpecificTravelReduction} ABOUT HERE ]} \\
+{[ FIGURE \ref{fig:ByGroupReduction_GlobalStrategy} ABOUT HERE ]} \\
 \end{center}
+
+% \vspace{0.5em}
+% \begin{center}
+% {[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
+% \end{center}
 
@@
 \begin{center}
-{[ FIGURE \ref{fig:GlobalTravelReduction} ABOUT HERE ]} \\
+{[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
 \end{center}
@@
     $g,g'$ & Reference group ($g$) and out-group $g'$ & Terena, Guarani; Bororó, Jaguapiru \\
+    \bottomrule
+  \end{tabular} 
+\end{table}
+
+\begin{table}[h]
+  \centering
+  \caption{\textbf{Independent and dependent analysis variables.} Reducing
+  mobility to the city (first row) hypothetically increases the reduction in
+cumulative cases in the Reserve (second row).}
+  \label{tab:AnalysisVariables}
+  \begin{tabular}{rp{3.0in}l} 
+    \toprule
+    variable & description & values \\
+    \midrule 
+    $\Delta \gamma$ and $\Delta \gamma_g$ & Total city mobility reduction and 
+      group-level mobility reduction & $0.0,0.2,\ldots,0.8$ \\
+      $\langle \Delta I^\mathrm{Tot} \rangle$ and $\langle \Delta I^\mathrm{Tot}_g
+      \rangle$  & Parameter-set ensemble averages of total and group-level
+      cumulative incidence reduction & [0.0, 1.0] \\
     \bottomrule
@@
 
+% \begin{figure}[H]
+
+% \caption{We hypothesize that overall reduction in disease incidence ($\Delta I$, y-axis)  depends on 
+% strategy to distribute the total aid to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
+% \label{fig:ByGroupReduction_GlobalStrategy}
+% \vspace{0.3em}
+% {\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByStrategy.pdf}
+
+% }
+% \end{figure}
+
+\begin{figure}[H]
+
+\caption{Disease reduction ($\Delta I$, y-axis) depends on the possibility of
+  re-infection and the
+distribution strategy to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
+\label{fig:ByReinfectionByStrategy}
+\vspace{0.3em}
+{\centering
+  \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByReinfectionByStrategy.pdf}
+
+}
+
+\end{figure}%
+
 \begin{figure}[H]
@@
 ($\Delta \gamma$, x-axis).}
-\label{fig:GlobalTravelReduction}
+\label{fig:ByGroupReduction_GlobalStrategy}
 \vspace{0.3em}
@@
 
-\begin{figure}[H]
-
-\caption{We hypothesize that overall reduction in disease incidence ($\Delta I$, y-axis)  depends on 
-strategy to distribute the total aid to reduce travel between Dourados Reserve and city ($\Delta \gamma$, x-axis).}
-\label{fig:GroupSpecificTravelReduction}
-\vspace{0.3em}
-{\centering \includegraphics[width=0.95\textwidth,height=\textheight]{Figures/ResultsSketch_ByStrategy.pdf}
-
-}
-
-\end{figure}%
-
 \end{document}

### 2024-07-26 ms.tex
going to write analysis sketch
added thought about equity and planning to disc

@@
 potential economic interventions that target aid to different groups to
-reduce infections in the reserve.
-
+reduce infections in the reserve. The effect of different amounts of aid to
+different groups 
 \vspace{0.5em}
@@
 for the city population are ignored.
+The $\gamma_g$ parameters are proxies for resources allocated to 
+economic aid or other programs to support to group $g$ to remain on the Reserve 
+during outbreaks.
 
@@
 \begin{center}
-{[ FIGURE \ref{fig:ByGroupReduction_GlobalStrategy} ABOUT HERE ]} \\
+{[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
 \end{center}
-
-% \vspace{0.5em}
-% \begin{center}
-% {[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
-% \end{center}
 
@@
 \begin{center}
-{[ FIGURE \ref{fig:ByReinfectionByStrategy} ABOUT HERE ]} \\
+{[ FIGURE \ref{fig:ByGroupReduction_GlobalStrategy} ABOUT HERE ]} \\
 \end{center}
@@
     health behaviors from others~\autocite{Turner2023a}.
+  \item
+    Currently we do not connect how much it costs to implement these programs,
+    and there may be difficult questions over what equality means: is it
+    spending the same amount of money per group or preventing the same number of
+    hypothetical infections for each group? Of course a scientific article 
+    cannot provide the answer, but here we do provide a program for evaluating
+    different choices to prototype planned responses to pandemic outbreaks and
+    reason quantitatively about alternative courses of action.
 \end{itemize}
@@
 \begin{figure}[H]
-\caption{Time series of COVID-19 spread through the Dourados Indigenous
-Reserve, with significant events highlighted. The first B.1.1 variant
+\caption{\textbf{Time series of COVID-19 spread through the Dourados Indigenous
+Reserve,} with significant events highlighted. The first B.1.1 variant
 cases were introduced by the Guarani ethnicity (1), followed about one
@@
 \caption{
-  We can model the relevant groups for mixing as either village
-residence or ethnicity. Mixing, and mobility parameters
-would be set accordingly. Mixing is asymmetric, with $c_{gg}$ representing the
+\textbf{Relevant groups for modeling inter-group mixing could be village
+residence or ethnicity.} Mixing is asymmetric, with $c_{gg}$ representing the
 contact rate within group $g$, and each $c_{gg'}$ representing
@@
 outgroup $g'$, e.g., $c_{GT}$ is the contact rate at which Guarani encounter
-members of the Terena. The $\gamma_g$ parameters would be reduced through
-economic aid to group $g$ for them to remain on the Reserve during
-outbreaks. \mt{Looks like some labels are cut off; will fix in later draft.}}
+members of the Terena (or vice-versa). \mt{Looks like some labels are cut off; will fix in later draft.}}
 \label{fig:ModelSketch}

### 2024-07-29 ms.tex
added ethnicity par stub in Intro; ready for analysis

@@
 \item
+  Ethnic identity is a complicated, contested concept due to coloinialism and
+  various waves of oppression, repression, and genocide against Indigenous
+  peoples in South America, which included various attempts at dismantling 
+  and renaming pre-colonial ethnic identity~\autocite{Brandão1990}.
+\item
   Second, we aim to use a parameterized model to understand how
@@
 
+To quantitatively compare hypothetical intervention strategies we analyze reduction
+in total Once implemented,
+
 \vspace{0.5em}
@@
 
-\caption{Disease reduction ($\Delta I$, y-axis) depends on the possibility of
+\caption{\textbf{Intervention efficacy depends on strategy and epidemiology.} Disease reduction ($\Delta I$, y-axis) depends on the possibility of
   re-infection and the
@@
 
-\caption{Global reduction in travel to and from the city could have
+\caption{\textbf{Intervention efficacy will vary between groups.} 
 Infection reduction ($\Delta I$, y-axis) for different groups could be
-different as global reduction in travel to and from the city decreases
+different under different mobility-reduction strategies
 ($\Delta \gamma$, x-axis).}

### 2024-07-29 ms.tex
more refs for ethnicity intro par

@@
   peoples in South America, which included various attempts at dismantling 
-  and renaming pre-colonial ethnic identity~\autocite{Brandão1990}.
+  and renaming pre-colonial ethnic
+  identity~\autocite{Brandão1990,FerreiraThomazDeAlmeida2018,Soratto2022}.
 \item

### 2024-07-30 ms.tex
started analysis sketch
updated some figs

@@
 
-To quantitatively compare hypothetical intervention strategies we analyze reduction
-in total Once implemented,
+To quantitatively compare hypothetical intervention strategies we will analyze 
+model outcomes in different contexts, modeled as different parameter settings.
+Since this document is for planning the study and analysis, here we present a sketch
+of results we plan to obtain and subsequent analysis of these results. In the first
+analysis, we examine how intervention efficacy varies with the possibility of
+re-infection and mobility reduction strategy
+(Figure~\ref{fig:ByReinfectionByStrategy}). In the second, we examine how
+intervention efficacy varies for different groups for a given mobility reduction
+strategy (Figure~\ref{fig:ByGroupReduction_GlobalStrategy}). These analyses 
+promise to account for some important sources of complexity in 
+the Dourados Reserve-city system that could affect intervention outcomes.
+
+\subsection{Aid strategy and epidemiological details affect incidence reduction}
+
+We hypothesize that reducing incidence of diseases or variants with 
+re-infection (e.g., Zeta COVID variant) may require greater investment 
+for the same reduction in incidence
+compared to diseases or variants without re-infection (e.g., B.1.1 variant).
+This would result in a sharper s-curve from zero reduction 
+to maximum reduction, with 50\% of maximum reduction occurring 
+at a greater value of $\Delta \gamma$, across aid strategies (illustrated in
+difference between orange, re-infection, and blue lines, no re-infection, in 
+Figure~\ref{fig:ByReinfectionByStrategy}).
+
+We also hypothesize that it will be more effective to target individuals based on
+geographic location compared to ethnic identity, assuming there are no significant
+differences in transmissibility, which may exist due to physical, medical or
+behavioral differences between ethnic groups \mt{we plan to return to this point in
+the Discussion}. This will result in efficacy curves that reach maximal
+effectiveness sooner (compare \emph{80-20} strategy with 
+80\% to Jaguapiru–dashed lines–to \emph{80-20}, with 80\% to Guarani (all)–dotted
+lines–and the \emph{Equal} strategy–solid line;
+Figure~\ref{fig:ByReinfectionByStrategy}).
 
@@
 \end{center}
+
+\subsection{Intervention efficacy varies by group}
 
@@
     spending the same amount of money per group or preventing the same number of
-    hypothetical infections for each group? Of course a scientific article 
+    hypothetical infections for each group? Should past discrimination and
+    government-imposed hardship be included in the calculus of developing
+    mobility-reduction strategies? Of course a scientific article 
     cannot provide the answer, but here we do provide a program for evaluating
